### 2020-07-16 Organismal_paper/Organismal_manuscript_and_code.Rmd
changes after CWD comments

---
bibliography: org_manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
---

@@

Siphonophores are free-living predatory colonial hydrozoan cnidarians found in every ocean of the world. Siphonophore tentilla (tentacle side branches) are unique biological structures for prey capture, composed of a complex arrangement of cnidocytes (stinging cells) bearing different types of nematocysts (stinging capsules) and auxiliary structures. Tentilla present an extensive morphological and functional diversity across species. While associations between tentilla form and diet have been reported, the evolutionary history giving rise to this morphological diversity is largely unexplored. Here we examine the evolutionary gains and losses of novel tentillum substructures and nematocyst types on the most recent siphonophore phylogeny. Tentilla have a precisely coordinated high-speed strike mechanism of synchronous unwinding and nematocysts discharge. Here we characterize the kinematic diversity of this prey capture reaction using high-speed video and find relationships with morphological characters. Since tentillum discharge occurs in synchrony across a broad morphological diversity, we evaluate how phenotypic integration is maintaining character correlations across evolutionary time. We found that the the tentillum morphospace has low dimensionality, we identified instances of heterochrony and morphological convergence, and generated hypotheses on the diets of understudied siphonophore species. Our findings indicate that siphonophore tentilla are phenotypically integrated structures with a complex evolutionary history leading to a phylogenetically structured diversity of forms which are predictive of kinematic performance and feeding habits.

@@

Siphonophores have fascinated zoologists for centuries for their extremely subspecialized colonial organization and integration. Today we have a comprehensive taxonomic coverage on the morphological diversity of this group due to the extensive work of siphonophore taxonomists in the past few decades  [@pugh1983benthic;@pugh1986new;@pugh1988two;@pugh2001review;@hissmann2005situ;@haddock2005re;@dunn2005marrus;@bardi2007taxonomic;@pugh2010three;@pugh2014review], which has been elegantly synthesized in detailed synopsis [@totton1965synopsis, @mapstone2014global]. In addition, recent advances in phylogenetic analyses of siphonophores [@munro2018improved;@damian2019evolution] have provided a macroevolutionary context to interpret this diversity. With these assets in hand, we can now begin to study siphonophores from an orthogonal perspective, focusing on the diversity and evolutionary history of specific structures. Here we focus on one of such structures: the tentilla. Like many cnidarians, siphonophore tentacles bear side branches (tentilla) with nematocysts. But unlike other cnidarians, most siphonophore tentilla are dynamic structures that react to prey encounters by rapidly unfolding the nematocyst battery to slap around the prey. This maximizes the surface area of contact between the nematocysts and the prey they fire upon. In addition, siphonophore tentilla present a remarkable diversity of morphologies, sizes, and nematocyst complements (Fig. \@ref(figure2)). Our overarching aim is to organize all this phenotypic diversity in a phylogenetic context, and identify the evolutionary processes that generated it

![\label{figure2} Tentillum diversity. The illustrations delineate the pedicle (green), involucrum (blue), cnidoband (orange), elastic strands (pink), terminal structures (yellow). Heteroneme nematocysts (stenoteles in C,E,F,G and mastigophores in H,I) are depicted in red for some species. A - *Erenna laciniata*, 10x. B - *Lychnagalma utricularia*, 10x. C - *Agalma elegans*, 10x. D - *Resomia ornicephala*, 10x. E - *Frillagalma vityazi*, 20x. F - *Bargmannia amoena*, 10x. G - *Cordagalma* sp., reproduced from Carré 1968. H - *Lilyopsis fluoracantha*, 20x. I - *Abylopsis tetragona*, 20x.](~/tentilla_morph/Figures_final/figure2.jpg)

@@

Nematocysts are unique biological weapons for defense and prey capture exclusive to the phylum Cnidaria. Mariscal ([@mariscal1974nematocysts]) reported that hydrozoans have the largest diversity of nematocyst types among cnidarians. Among them, siphonophores present the greatest variety of types [@mapstone2014global], and vary widely across taxa in which and how many types they carry on their tentacles. Werner [@werner1965nesselkapseln] noted that there are nine types of nematocyst found in siphonophores, of which four, anacrophore rhopalonemes, acrophore rhopalonemes, homotrichous anisorhizas, and birhopaloids, are unique to them. Heteroneme and haploneme nematocysts serve penetrant and entangling functions, while rhopalonemes and desmonemes work by adhering to the surface of the prey. While recent descriptive studies have expanded and confirmed our understanding of this diversity, the evolutionary history of nematocyst type gain and loss in siphonophores remains unexplored. Thus, here we reconstruct the evolution of shifts, gains, and losses of nematocyst types, subtypes, and other major categorical traits that led to the extant diversity we see in siphonophore tentilla.

@@

![\label{figure7} Siphonophore cladogram with the main categorical character gains (green) and losses (red) mapped. Some branch lengths were modified from the Bayesian chronogram to improve readability. The main visually distinguishable tentillum types are sketched next to the species that bear them, showing the location and arrangement of the main characters. In large, complex-shaped euphysonect tentilla, haplonemes were omitted for simplification. The hypothesized phylogenetic placement of the rhizophysid *Bathyphysa conifera*, for which no molecular data are yet available, was added manually (dashed line).](Figures_final/figure7.jpg)

*Evolution of tentillum and nematocyst characters* – Most (74%) characters present a significant phylogenetic signal, yet only total nematocyst volume, haploneme length, and heteroneme-to-cnidoband length ratio had a phylogenetic signal with K larger than 1. Total nematocyst volume and cnidoband-to-heteroneme length ratio showed strongly conserved phylogenetic signals. The majority (67%) of characters were best fitted by BM models, indicating a history of neutral constant divergence. We did not find any relationship between phylogenetic signal and BM model support. One-third of the characters measured in [@damian2019evolution] support a non-phylogenetic generative model, indicating they are not phylogenetically conserved (SM10). Haploneme nematocyst length was the only character with support for an EB model of decreasing rate of evolution with time. No character had support for a single-optimum OU model (when not informed by feeding guild regime priors). The model adequacy tests (SM11) indicate that many characters may have a relationship between the states and the rates of evolution (Sasr) not captured in the basic models compared here, accompanied by a signal of unaccounted rate heterogeneity (Cvar). No characters show significant deviations in the overall rate of evolution estimated (Msig). Some characters show a perfect fit (no significant deviations across all metrics) under BM evolution, such as heteroneme shape, length, width & volume, haploneme width & SA/V, tentacle width and pedicle width. Haploneme row number and rhopaloneme shape have significant deviations across four metrics, indicating that BM (best model) is a poor fit. These characters likely evolved under complex models which would require many more data points than we have available to fit with accuracy.

@@

*Evolution of nematocyst shape* – The greatest evolutionary change in haploneme nematocyst shape occurred in a single shift towards elongation in the stem of Tendiculophora, which contains the majority of described siphonophore species other than Cystonects, *Apolemia*, and Pyrostephidae. There is one secondary return to more oval, less elongated haplonemes in *Erenna*, but it does not reach the sphericity present in Cystonectae or Pyrostephidae (Fig. \@ref(figure10)). Heteroneme evolution presents a less discrete evolutionary history, where Tendiculophora evolved more elongate heteronemes, but the difference between theirs and other siphonophores is much smaller than the variation in shape within Tendiculophora, bearing no phylogenetic signal. In this clade, the evolution of heteroneme shape has diverged in both directions, and there is no correlation with haploneme shape (Fig. \@ref(figure10)), which has remained fairly constant (elongation between 1.5 and 2.5). Haploneme and heteroneme shape share 21% of their variance across extant values, and 53% of the variance in their shifts along the branches of the phylogeny. However, much of this correlation is due to the sharp contrast between Pyrostephidae and their sister group Tendiculophora. We searched for regime shifts in the evolution of haploneme nematocyst shape characters using SURFACE [@ingram2013surface]. SURFACE identified eight distinct OU regimes in the evolutionary history of haploneme length and width (Fig. \@ref(surface)A). The different regimes are located (1) in cystonects, (2) in most of Tendiculophora, (3) in most diphyomorphs, (4) in *Cordagalma ordinatum*, (5) in *Stephanomia amphytridis*, (6) in pyrostephids, (7) in *Diphyes dispar* + *Abylopsis tetragona*, and (8) in *Erenna* spp.

@@

*Morphospace occupation* – In order to examine the occupation structure of the morphospace across all siphonophore species in the dataset, we cast a PCA on the data after transforming inapplicable states (due to absence of character) to zeroes. This allows us to accommodate species with many missing characters (such as cystonects or apolemiids), and to account for common absences as morphological similarities. In this ordination, PC1 explains 47.45% of variation (aligned with cnidoband size) and PC2 explains 16.73% of variation (aligned with heteroneme volume and involucrum). When superimposing feeding guilds onto the morphospace (Fig. \@ref(morphodiet)), we find that the morphospaces of each feeding guild are only slightly overlapping in PC1 and PC2. A phylogenetic MANOVA showed that feeding guilds explain 27.63% of variance across extant species (p value < 0.000001), and 20.97% of the variance in the tree species accounting for phylogeny, an outcome significantly distinct from the expectation under neutral evolution (p-value = 0.0196). In addition, a morphological disparity analysis accounting for phylogenetic structure shows that the morphospace of fish specialists is significantly broader than that of generalists and other specialists. This is due to the large morphological disparities between cystonects and piscivorous euphysonects. There are no significant differences among the morphospace disparities of the other feeding guilds.

@@

*Convergent evolution* – Convergence is a widespread evolutionary phenomenon where distantly related clades independently evolve similar phenotypes. Using the package SURFACE [@ingram2013surface], we identified convergence in haploneme nematocyst shape and in morphospace position. In [@damian2019evolution], we identified haploneme nematocyst shape as one of the traits associated with the convergent evolution of piscivory. Here we find that indeed wider haploneme nematocysts have convergently evolved in the piscivore cytonects and *Erenna* spp. (Fig. \@ref(surface)A). Extremely narrow haplonemes have also evolved convergently in *Cordagalma ordinatum* and copepod specialist calycophorans such as *Sphaeronectes koellikeri*. When integrating many traits into a couple principal components, we find two distinct convergences between euphysonects and calycophorans with a reduced prey capture apparatus. Those convergences are between *Frillagalma vityazi* and calycophorans, and once again between *Cordagalma ordinatum* and *Spaheronectes koellikeri* (Fig. \@ref(surface)B).

*Functional morphology of tentillum and nematocyst discharge* –  Tentillum and nematocyst discharge high speed measurements are available in the Dryad repository. While the sample sizes of these measurements were insufficient to draw reliable statistical results at a phylogenetic level, we did observe patterns that may be relevant to their functional morphology. For example, cnidoband length is strongly correlated with discharge speed (p value = 0.0002). This is probably the sole driver of the considerable difference between euphysonect and calycophoran tentilla discharge speeds (average discharge speeds: 225.0mm/s and 41.8mm/s respectively; t-test p value = 0.011), since the euphysonects have larger tentilla than the calycophorans among the species recorded. In addition, we observed that calycophoran haploneme tubules fire faster than those of euphysonects (T-test p value = 0.001). Haploneme nematocysts discharge 2.8x faster than heteroneme nematocysts (T-test p value = 0.0012). Finally, we observed that the stenoteles of the Euphysonectae discharge a helical filament that “drills” itself through the medium it penetrates as it everts.

@@

*On the evolution of tentilla morphology* -- The evolutionary rate covariance results in [@damian2019evolution] indicate that tentilla are not only phenotypically integrated but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. The rate covariance results are congruent with the evolutionary correlations we found (Fig. \@ref(figure8)). In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos (pers. obs.) show that the involucrum helps direct the whiplash of the uncoiling cnidoband distally (towards the prey). The clade Tendiculophora contains far more species than its relatives Cystonectae, Apolemiidae, and Pyrostephidae. An increase in clade richness and ecological diversification can be triggered by a ‘key innovation’ [@simpson1955major]. The evolutionary innovation of the Tendiculophora tentilla with shooting cnidobands and modular regions may have facilitated further dietary diversification. A specific instance of this may have been the access to the abundant small crustacean prey such as copepods. The rapid darting escape response of copepods may preclude their capture in siphonophores without shooting cnidobands.

The siphonophore tentillum morphospace has a fairly low extant dimensionality due to having an evolutionary history with many synchronous, correlated changes. This is consistent with strong phenotypic integration where genetic and developmental correlations are maintained by natural selection to preserve a complex function across the wide variety of morphologies present. Since most tentillum characters develop from a common bud (budding tentilla near the base of the tentacle), structural correlations are expected. Similarly, correlations between the features of different nematocyst subtypes within a species are also expected given their common evolutionary and developmental origin [@bode1988control;@david2008evolution]. However, we also found correlations between nematocyst and tentillum characters. Siphonophore tentacle nematocysts (in their cnidocytes) are not produced nor matured in the developing tentillum. These cnidocytes are produced by dividing cnidoblasts in the basigaster (basal swelling of the gastrozooid). Once the cnidocytes have assembled the nematocyst, they migrate outward along the tentacle [@carre1972study] and position themselves in the tentillum according to their type and size [@skaer1988formation]. Thus, the developmental programs that produce the observed nematocyst morphologies are spatially separated from those producing the tentillum morphologies. Therefore, we hypothesize the genetic correlations and phenotypic integration between tentillum and nematocyst characters are maintained through natural selection on separate regulatory networks, out of the necessity to work together and meet the spatial, mechanical, and functional constraints of their prey capture behavior.

>>>>

---
bibliography: org_manuscript.bib
csl: pnas.csl
site: "bookdown::bookdown_site"
output:
  bookdown::pdf_document2:
    toc: FALSE
indent: TRUE
monofont: "Times New Roman"
fontsize: 12pt
header-includes:
  - \usepackage[left]{lineno}
  - \linenumbers
  - \usepackage{setspace}
  - \doublespacing
  - \usepackage{placeins}
---

@@

Siphonophores are free-living predatory colonial hydrozoan cnidarians found in every ocean of the world. Siphonophore tentilla (tentacle side branches) are unique biological structures for prey capture, composed of a complex arrangement of cnidocytes (stinging cells) bearing different types of nematocysts (stinging capsules) and auxiliary structures. Tentilla present an extensive morphological and functional diversity across species. While associations between tentilla form and diet have been reported, the evolutionary history giving rise to this morphological diversity is largely unexplored. Here we examine the evolutionary gains and losses of novel tentillum substructures and nematocyst types on the most recent siphonophore phylogeny. Tentilla have a precisely coordinated high-speed strike mechanism of synchronous unwinding and nematocysts discharge. Here we characterize the kinematic diversity of this prey capture reaction using high-speed video and find relationships with morphological characters. Since tentillum discharge occurs in synchrony across a broad morphological diversity, we evaluate how phenotypic integration is maintaining character correlations across evolutionary time. We found that the tentillum morphospace has low dimensionality, we identified instances of heterochrony and morphological convergence, and generated hypotheses on the diets of understudied siphonophore species. Our findings indicate that siphonophore tentilla are phenotypically integrated structures with a complex evolutionary history leading to a phylogenetically structured diversity of forms which are predictive of kinematic performance and feeding habits.

@@

Siphonophores have fascinated zoologists for centuries for their extremely subspecialized colonial organization and integration. Today we have a comprehensive taxonomic coverage on the morphological diversity of this group due to the extensive work of siphonophore taxonomists in the past few decades  [@pugh1983benthic;@pugh1986new;@pugh1988two;@pugh2001review;@hissmann2005situ;@haddock2005re;@dunn2005marrus;@bardi2007taxonomic;@pugh2010three;@pugh2014review], which has been elegantly synthesized in detailed synopsis [@totton1965synopsis, @mapstone2014global]. In addition, recent advances in phylogenetic analyses of siphonophores [@munro2018improved;@damian2019evolution] have provided a macroevolutionary context to interpret this diversity. With these assets in hand, we can now begin to study siphonophores from an orthogonal perspective, focusing on the diversity and evolutionary history of specific structures. Here we focus on one of such structures: the tentilla. Like many cnidarians, siphonophore tentacles bear side branches (tentilla) with nematocysts (Fig. \@ref(anatomy)). But unlike other cnidarians, most siphonophore tentilla are dynamic structures that react to prey encounters by rapidly unfolding the nematocyst battery to slap around the prey. This maximizes the surface area of contact between the nematocysts and the prey they fire upon. In addition, siphonophore tentilla present a remarkable diversity of morphologies (Fig. \@ref(figure2)), sizes, and nematocyst complements (Fig. \@ref(nematocysts)). Our overarching aim is to organize all this phenotypic diversity in a phylogenetic context, and identify the evolutionary processes that generated it.

![\label{anatomy} Siphonophore anatomy. A - Nanomia sp. siphonophore colony (photo by Catriona Munro). B, C - Illustration of a Nanomia colony, gastrozooid, and tentacle closeup (by Freya Goetz). D - Nanomia sp. Tentillum illustration and main parts. E - Differential interference contrast micrograph of the tentillum illustrated in D. Figure reproduced from Damian-Serrano et al. 2020 with permission.](~/tentilla_morph/Organismal_paper/figures/anatomy.png)

@@

![\label{figure2} Tentillum diversity. The illustrations delineate the pedicle, involucrum, cnidoband, elastic strands, terminal structures. Heteroneme nematocysts (stenoteles in C,E,F,G and mastigophores in H,I) are only depicted for some species. A - *Erenna laciniata*, 10x. B - *Lychnagalma utricularia*, 10x. C - *Agalma elegans*, 10x. D - *Resomia ornicephala*, 10x. E - *Frillagalma vityazi*, 20x. F - *Bargmannia amoena*, 10x. G - *Cordagalma* sp., reproduced from Carré 1968. H - *Lilyopsis fluoracantha*, 20x. I - *Abylopsis tetragona*, 20x.](~/tentilla_morph/Organismal_paper/figures/diversity.png)

Nematocysts are unique biological weapons for defense and prey capture exclusive to the phylum Cnidaria. Mariscal ([@mariscal1974nematocysts]) reported that hydrozoans have the largest diversity of nematocyst types among cnidarians. Among them, siphonophores present the greatest variety of types [@mapstone2014global], and vary widely across taxa in which and how many types they carry on their tentacles (Fig. \@ref(nematocysts)). Werner [@werner1965nesselkapseln] noted that there are nine types of nematocyst found in siphonophores, of which four, anacrophore rhopalonemes, acrophore rhopalonemes, homotrichous anisorhizas, and birhopaloids, are unique to them. Heteroneme and haploneme nematocysts serve penetrant and entangling functions, while rhopalonemes and desmonemes work by adhering to the surface of the prey. While recent descriptive studies have expanded and confirmed our understanding of this diversity, the evolutionary history of nematocyst type gain and loss in siphonophores remains unexplored. Thus, here we reconstruct the evolution of shifts, gains, and losses of nematocyst types, subtypes, and other major categorical traits that led to the extant diversity we see in siphonophore tentilla.

![\label{nematocysts} Phylogenetic distribution of nematocyst types, subtypes, functions, and locations in the zooid across the major siphonophore clades. Illustrations reproduced with permission from Mapstone 2014. Undischarged capsules to the left, discharged to the right. Agalmatidae* here refers only to the genera *Agalma*, *Athorybia*, *Halistemma*, and *Nanomia*. ](~/tentilla_morph/Organismal_paper/figures/nematocysts.png)

Distantly related organisms that evolved to feed on similar resources often evolve similar adaptations [@winemiller2015functional]. In [@damian2019evolution], we found strong associations between piscivory and haploneme shape across distantly related siphonophore lineages. These associations could have been produced by convergent changes in the adaptive optima of these characters. Here we set out to test this hypothesis using comparative model fitting methods. Analyzing the diversity of morphological states from a phylogenetic perspective allows us to identify the specific evolutionary processes that gave rise to it. Here we fit and compare a variety of macroevolutionary models to siphonophore tentilla morphology measurement data to identify instances of neutral divergence, stabilizing selection, changes in the speed of evolution, convergent evolution, and paedomorphosis (larval characters present in mature colonies).

@@

![\label{figure7} Siphonophore cladogram with the main categorical character gains (green) and losses (red) mapped. Some branch lengths were modified from the Bayesian chronogram to improve readability. The main visually distinguishable tentillum types are sketched next to the species that bear them, showing the location and arrangement of the main characters. In large, complex-shaped euphysonect tentilla, haplonemes were omitted for simplification. The hypothesized phylogenetic placement of the rhizophysid *Bathyphysa conifera*, for which no molecular data are yet available, was added manually (dashed line).](~/tentilla_morph/Figures_final/figure7.jpg)

*Evolution of tentillum and nematocyst characters* – Most (74%) characters present a significant phylogenetic signal, yet only total nematocyst volume, haploneme length, and heteroneme-to-cnidoband length ratio had a phylogenetic signal with K larger than 1. Total nematocyst volume and cnidoband-to-heteroneme length ratio showed strongly conserved phylogenetic signals. The majority (67%) of characters were best fitted by BM models, indicating a history of neutral constant divergence. We did not find any relationship between phylogenetic signal and specific model support, where characters with high and low phylogenetic signal were broadly distributed among the best fitted for each model. One-third of the characters measured in [@damian2019evolution] did not recover significant support for any of the phylogenetic models tested, indicating they are either not phylogenetically conserved, or they evolved under a complex evolutionary process not represented among the models tested (SM10). Haploneme nematocyst length was the only character with support for an EB model of decreasing rate of evolution with time. No character had support for a single-optimum OU model (when not informed by feeding guild regime priors). The model adequacy tests (SM11) indicate that many characters may have a relationship between the states and the rates of evolution (Sasr) not captured in the basic models compared here, accompanied by a signal of unaccounted rate heterogeneity (Cvar). No characters show significant deviations in the overall rate of evolution estimated (Msig). Some characters show a perfect fit (no significant deviations across all metrics) under BM evolution, such as heteroneme shape, length, width & volume, haploneme width & SA/V, tentacle width and pedicle width. Haploneme row number and rhopaloneme shape have significant deviations across four metrics, indicating that BM (best model) is a poor fit. These characters likely evolved under complex models which would require many more data points than we have available to fit with accuracy.

@@

*Evolution of nematocyst shape* – The greatest evolutionary change in haploneme nematocyst shape occurred in a single shift towards elongation in the stem of Tendiculophora, which contains the majority of described siphonophore species, *i.e.* all siphonophores other than Cystonects, *Apolemia*, and Pyrostephidae. There is one secondary return to more oval, less elongated haplonemes in *Erenna*, but it does not reach the sphericity present in Cystonectae or Pyrostephidae (Fig. \@ref(figure10)). Heteroneme evolution presents a less discrete evolutionary history. Tendiculophora evolved more elongate heteronemes along the stem, but the difference between theirs and other siphonophores’ is much smaller than the variation in shape within Tendiculophora, bearing no phylogenetic signal within this clade. In this clade, the evolution of heteroneme shape has diverged in both directions, and there is no correlation with haploneme shape (Fig. \@ref(figure10)), which has remained fairly constant (elongation between 1.5 and 2.5).

Haploneme and heteroneme shape share 21% of their variance across extant values, and 53% of the variance in their shifts along the branches of the phylogeny. However, much of this correlation is due to the sharp contrast between Pyrostephidae and their sister group Tendiculophora. We searched for regime shifts in the evolution of haploneme nematocyst shape characters using SURFACE [@ingram2013surface]. SURFACE identified eight distinct OU regimes in the evolutionary history of haploneme length and width (Fig. \@ref(surface)A). The different regimes are located (1) in cystonects, (2) in most of Tendiculophora, (3) in most diphyomorphs, (4) in *Cordagalma ordinatum*, (5) in *Stephanomia amphytridis*, (6) in pyrostephids, (7) in *Diphyes dispar* + *Abylopsis tetragona*, and (8) in *Erenna* spp.

@@

*Morphospace occupation* – In order to examine the occupation structure of the morphospace across all siphonophore species in the dataset, we cast a PCA on the data after transforming inapplicable states (due to absence of character) to zeroes. This allows us to accommodate species with many missing characters (such as cystonects or apolemiids), and to account for common absences as morphological similarities. In this ordination, PC1 (aligned with cnidoband size) explains 47.45% of variation and PC2 (aligned with heteroneme volume and involucrum length) explains 16.73% of variation . When superimposing feeding guilds onto the morphospace (Fig. \@ref(morphodiet)), we find that the morphospaces of each feeding guild are only slightly overlapping in PC1 and PC2. A phylogenetic MANOVA showed that feeding guilds explain 27.63% of variance across extant species (p value < 0.000001), and 20.97% of the variance in the tree species accounting for phylogeny, an outcome significantly distinct from the expectation under neutral evolution (p-value = 0.0196). In addition, a morphological disparity analysis accounting for phylogenetic structure shows that the morphospace of fish specialists is significantly broader than that of generalists and other specialists. This is due to the large morphological disparities between cystonects and piscivorous euphysonects. There are no significant differences among the morphospace disparities of the other feeding guilds.

@@

*Convergent evolution* – Convergence is a widespread evolutionary phenomenon where distantly related clades independently evolve similar phenotypes. When the dimensionality of the state space is small as it is in tentilla morphology, convergence is more likely given the same amount of evolutionary change. Using the package SURFACE [@ingram2013surface], we identified convergence in haploneme nematocyst shape and in morphospace position. In [@damian2019evolution], we identified haploneme nematocyst shape as one of the traits associated with the convergent evolution of piscivory. Here we find that indeed wider haploneme nematocysts have convergently evolved in the piscivore cytonects and *Erenna* spp. (Fig. \@ref(surface)A). Extremely narrow haplonemes have also evolved convergently in *Cordagalma ordinatum* and copepod specialist calycophorans such as *Sphaeronectes koellikeri*. When integrating many traits into a couple principal components, we find two distinct convergences between euphysonects and calycophorans with a reduced prey capture apparatus. Those convergences are between *Frillagalma vityazi* and calycophorans, and once again between *Cordagalma ordinatum* and *Spaheronectes koellikeri* (Fig. \@ref(surface)B).

*Functional morphology of tentillum and nematocyst discharge* –  Tentillum and nematocyst discharge high speed videos and measurements are available in the Dryad repository. While the sample sizes of these measurements were insufficient to draw reliable statistical results at a phylogenetic level, we did observe patterns that may be relevant to their functional morphology. For example, cnidoband length is strongly correlated with discharge speed (p value = 0.0002). This explains much  of the considerable difference between euphysonect and calycophoran tentilla discharge speeds (average discharge speeds: 225.0mm/s and 41.8mm/s respectively; t-test p value = 0.011), since the euphysonects have larger tentilla than the calycophorans among the species recorded. In addition, we observed that calycophoran haploneme tubules fire faster than those of euphysonects (t-test p value = 0.001). Haploneme nematocysts discharge 2.8x faster than heteroneme nematocysts (t-test p value = 0.0012). Finally, we observed that the stenoteles of the Euphysonectae discharge a helical filament that “drills” itself through the medium it penetrates as it everts.

@@

*On the evolution of tentilla morphology* -- The evolutionary rate covariance results in [@damian2019evolution] indicate that tentilla are not only phenotypically integrated (with widespread evolutionary correlations across structures) but also show patterns of evolutionary modularity, where different sets of characters appear to evolve in stronger correlations among each other than with other characters [@wagner1996homologues]. This may be indicative of the underlying genetic and developmental dependencies among closely homologous nematocyst types (such as desmonemes and rhopalonemes) and structures. The rate covariance results are congruent with the evolutionary correlations we found (Fig. \@ref(figure8)). In addition, these evolutionary modules point to hypothetical functional modules. For example, the coiling degree of the cnidoband and the extent of the involucrum have correlated rates of evolution, while high-speed videos show that the involucrum helps direct the whiplash of the uncoiling cnidoband distally (towards the prey). The clade Tendiculophora contains far more species than its relatives Cystonectae, Apolemiidae, and Pyrostephidae. An increase in clade richness and ecological diversification can be triggered by a ‘key innovation’ [@simpson1955major]. The evolutionary innovation of the Tendiculophora tentilla with shooting cnidobands and modular regions may have facilitated further dietary diversification. A specific instance of this may have been the access to the abundant small crustacean prey such as copepods. The rapid darting escape response of copepods may preclude their capture in siphonophores without shooting cnidobands.

Siphonophore tentilla are beautifully complex and highly diverse. Our analyses show, however, that the siphonophore tentillum morphospace actually has a fairly low extant dimensionality due to having an evolutionary history with many synchronous, correlated changes. This can be due to many causes including structural constraints, shared developmental programs (by homology or proximity), or natural selection on separate developmental programs to preserve an adaptive function. A familiar example of functionally-maintained correlations between distinct structures can be found in maple trees, where leaf size, twig width, and branch spacing evolve in correlation to ensure that hydraulic supply and demand are balanced (Corner’s rules) [@ackerly1998leaf]. We believe some of the correlations we found are also maintained by natural selection to preserve a complex function across a wide variety of shapes and sizes. Most tentillum characters develop from a common bud (budding tentilla near the base of the tentacle),  and different nematocyst subtypes share a common evolutionary and developmental origin [@bode1988control;@david2008evolution]. In this sense, correlations within these modules are expected for reasons other than adaptive natural selection. However, we also found correlations between nematocyst and tentillum characters. Siphonophore tentacle nematocysts (in their cnidocytes) are not produced nor matured in the developing tentillum. These cnidocytes are produced by dividing cnidoblasts in the basigaster (basal swelling of the gastrozooid). Once the cnidocytes have assembled the nematocyst, they migrate outward along the tentacle [@carre1972study] and position themselves in the tentillum according to their type and size [@skaer1988formation]. Thus, the developmental programs that produce the observed nematocyst morphologies are spatially separated from those producing the tentillum morphologies, and they originate from evolutionarily distinct cell lineages. Therefore, we hypothesize the genetic correlations and phenotypic integration between tentillum and nematocyst characters are maintained through natural selection on separate regulatory networks, out of the necessity to work together and meet the spatial, mechanical, and functional constraints of their prey capture behavior. In order to adequately test these hypotheses, future work would need to study the genetic mechanisms underlying the development of tentilla from a comparative, evolutionary approach. Fortunately, the unique biology of siphonophore tentacles displays the full developmental sequence of tentilla along each tentacle, making siphonophores an ideal system for the comparative study of development.
